Unigram entropy in childes

### 2018-08-22 Unigram Positional Entropy.Rmd
Thoroughly knitted

```{r}
##functions
compile_child_utterances <- function(age_months) {
	# Given an age in months (integer), returns the child utterances from the Eng-NA corpus
	# for that age.
	d = get_utterances(collection="Eng-NA", age=age_months, role="Target_Child")
	return(d$gloss)
}

compile_adult_utterances <- function(age_months, adult_role) {
	# Given an age in months (integer), returns the adult_role utterances from the Eng-NA corpus
	# for that age (of the child to which the adult is speaking).
	# Valid are "Mother", "Investigator", "Sister", "Nurse", "Doctor" and "Father". 
	d = get_utterances(collection="Eng-NA", age=age_months, role=adult_role)
	return(d$gloss)
}

#IF WE CAN'T USE AN AGE RANGE
compile_ages <- function(age_range) {
	rv <- c()
	for (age in c(age_range[1]:age_range[2])) {
		rv <- c(rv, compile_child_utterances(age))
	}
	return(rv)
}
#IF WE CAN'T USE AN AGE RANGE
```

Get utterances
```{r get_utterances}
prov_utterances <- get_utterances(corpus = "Providence")

brown_utterances <- get_utterances(corpus = "Brown")

NA_utterances <- get_utterances(collection="Eng-NA")
```

compute estimates
```{r calc_pos_entropy}
get_entropies <- function(role, sen_length, utterances) {

@@

```{r ngram_frequencies}
m_prov <- prov_utterances %>%
          filter(speaker_role == "Target_Child") %>%
          mutate(length = str_count(stem, " ") + 1) %>% 
          filter(length >= 2)

@@

```{r Spanish unigram frequencies}
shiro_utterances <- get_utterances(language = "SPA", corpus="Shiro") 
#Venezuelan children narrative (113 children; half 1st graders half 4th graders; half high SES half low SES)

wagner_utterances <- get_utterances(corpus = "Wagner") #German corpus of children during daily routines (age range from 1,5 to 14,10)

@@

get_noneng_entropies <- function(role, sen_length, utterances) {

@@

entropies <- map(2:10, ~get_noneng_entropies("not child", .x, okayama_utterances)) %>%
  bind_rows() %>% filter(word_order <= length)

entropies_child <- map(2:10, ~ get_noneng_entropies("Target_Child", .x, okayama_utterances)) %>%
  bind_rows() %>% filter(word_order <= length)

@@

ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper, color = role)) +
  facet_wrap(~ length) + 
  geom_pointrange(position = position_dodge(.25)) +
  geom_smooth(se = F) +
  ggtitle("Child and Adult Positional Entropy vs. Word Position (Japanese Okayama corpus)")
```

>>>>

Get utterances
```{r get_utterances}
prov_utterances <- get_utterances(corpus = "Providence")
```

Compute Providence Corpus Estimates
```{r calc_pos_entropy}
get_entropies <- function(role, sen_length, utterances) {

@@

Compile all bigrams in corpus (save for another time) 
```{r ngram_frequencies}
m_prov <- prov_utterances %>%
          filter(speaker_role == "Target_Child") %>%
          mutate(length = str_count(stem, " ") + 1) %>% 
          filter(length >= 2)

@@

Non-English Unigram frequencies
```{r Spanish unigram frequencies}
shiro_utterances <- get_utterances(language = "SPA", corpus="Shiro") 
#Venezuelan children narrative (113 children; half 1st graders half 4th graders; half high SES half low SES)

@@

get_noneng_entropies <- function(role, sen_length, utterances, lang) {

  if(role == "Target_Child") {
    sub_utterances <- utterances %>%
      filter(speaker_role == "Target_Child") %>%
      mutate(stems = wordStem(gloss, language = lang)) %>% 
      mutate(length = str_count(stems, " ") + 1) 
  } else {
     sub_utterances <- utterances %>%
      filter(speaker_role != "Target_Child") %>%
      mutate(stems = wordStem(gloss, language = lang)) %>% 
      mutate(length = str_count(stems, " ") + 1)
  }

  tokens <- sub_utterances %>%
    filter(length == sen_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, stems) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n())

entropies <- map(2:10, ~get_noneng_entropies("not child", .x, shiro_utterances, "spanish")) %>%
  bind_rows() %>% filter(word_order <= length) #some utterances have ""

ggplot(entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F) +
  ggtitle("Adult Unigram Entropy vs. Word Position (Spanish Shiro Corpus with stemming)")

entropies_child <- map(2:10, ~ get_noneng_entropies("Target_Child", .x, shiro_utterances, "spanish")) %>%
  bind_rows() %>% filter(word_order <= length)

ggplot(entropies_child, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F) +
  ggtitle("Child Unigram Entropy vs. Word Position (Spanish Shiro Corpus with stemming)")

ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper, color = role)) +
  facet_wrap(~ length) + 
  geom_pointrange(position = position_dodge(.25)) +
  geom_smooth(se = F) +
  ggtitle("Adult and Child Unigram Entropy vs. Word Position (Spanish Shiro Corpus with stemming)")
```

```{r German Unigram Frequencies}
wagner_utterances <- get_utterances(corpus = "Wagner") #German corpus of children during daily routines (age range from 1,5 to 14,10)

entropies <- map(2:10, ~get_noneng_entropies("not child", .x, wagner_utterances, "german")) %>%
  bind_rows() %>% filter(word_order <= length) #some utterances have ""

ggplot(entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F) +
  ggtitle("Adult Unigram Entropy vs. Word Position (German Wagner Corpus with stemming)")

entropies_child <- map(2:10, ~ get_noneng_entropies("Target_Child", .x, wagner_utterances, "german")) %>%
  bind_rows() %>% filter(word_order <= length)

ggplot(entropies_child, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F) +
  ggtitle("Child Unigram Entropy vs. Word Position (German Wagner Corpus with stemming)")

ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper, color = role)) +
  facet_wrap(~ length) + 
  geom_pointrange(position = position_dodge(.25)) +
  geom_smooth(se = F) +
  ggtitle("Adult and Child Unigram Entropy vs. Word Position (German Wagner Corpus with stemming)")
```

```{r Japanese Unigram Entropy}
get_japanese_entropies <- function(role, sen_length, utterances) {

@@

entropies <- map(2:10, ~get_japanese_entropies("not child", .x, okayama_utterances)) %>%
  bind_rows() %>% filter(word_order <= length) #some utterances have "" counted for word_order

ggplot(entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F) +
  ggtitle("Adult Unigram Entropy vs. Word Position (Japanese Okayama Corpus)")

entropies_child <- map(2:10, ~ get_japanese_entropies("Target_Child", .x, okayama_utterances)) %>%
  bind_rows() %>% filter(word_order <= length)

ggplot(entropies_child, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F) +
  ggtitle("Child Unigram Entropy vs. Word Position (Japanese Okayama Corpus)")

@@

ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper, color = role)) +
  facet_wrap(~ length) + 
  geom_pointrange(position = position_dodge(.25)) +
  geom_smooth(se = F) +
  ggtitle("Adult and Child Unigram Entropy vs. Word Position (Japanese Okayama Corpus)")
```
